---
title: "Scripting task configuration files"
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction

The `tasks.json` configuration file is a the source of truth for validating a
hub and is also one of the most complex aspects of a hub. This configuration
file is a formal representation of an official collaborative forecasting
challenge that allows direct and unambiguous evaluation and comparison of model
outputs. It allows hub administrators to set a clear standard for model
submissions.

Creation of this configuration file is not easy, however---and it should not
be easy because setting a hub up for success requires a lot of care in
understanding what you are attempting to model. This includes knowledge of
what factors are important as modeling parameters and what is the best
representation of model outputs that allows for effective evaluation and
summarization so decisions can be made quickly.

While we provide [a hubTemplate](https://github.com/hubverse-org/hubTemplate/)
for use to get started with setting up a hub, JSON files are easy for a computer
to read, but not so easy for a human to read and less so to write. The hard part
of creating a `tasks.json` file should be what model parameters to use, not
constantly looking at JSON syntax or hunting down a missing comma.
To make this process easier, **we provide the `create_*()` family of functions
for creating `tasks.json` files programmatically**.  These can be useful for
creating your initial `task.json` file, amending an existing file's
configurations or even appending new rounds programmatically. By making use of
this functionality, hubs can even automate the process of updating their files
through Github Actions.  For example, the [variant nowcast
hub](https://github.com/reichlab/variant-nowcast-hub/tree/main) uses a [GitHub
Action
workflow](https://github.com/reichlab/variant-nowcast-hub/blob/29a9265d0b30f385fc47195d3d7f65ecda7a41d2/.github/workflows/create-modeling-round.yaml#L47)
that [runs a
script](https://github.com/reichlab/variant-nowcast-hub/blob/main/src/make_round_config.R)
to append a new round with updated variants to model on a weekly schedule!

In this vignette, we will use the config creation functions in `hubAdmin` to
create a `tasks.json` file that represents two rounds of a modeling effort to
predict influenza hospitalizations and optionally deaths 1 to 4 weeks ahead
with mean, median, and quantile predictions in the US, and optionally 5 states.

Before getting to the scripting, it's important to get a general look at what
a `tasks.json` file looks like structurally.

## Structure of a `tasks.json` configuration file

The structure of a `tasks.json` configuration file roughly looks like the
diagram below:

```
tasks.json
├─schema_version: "https://.../v4.0.0/tasks-schema.json"
├─rounds:
│ ├─model_tasks:
│ │   ├─task_ids:
│ │   │ └─[...]
│ │   ├─output_type:
│ │   │ └─[...]
│ │   └─target_metadata:
│ │     └─[...]
│ ├─round_id_from_variable: true
│ ├─round_id: origin_date
│ └─submissions_due:
│   └─[...]
├─output_type_id_datatype: "auto"
└─derived_task_ids:
  └─[...]
```

The first property in a `tasks.json` file is the `schema_version`, which provides
a URL to a schema that validates the file is in the expected structure.
The next property defined are the **rounds** which set the schedule and expected contents of model submissions
These rounds contain one or more [**modeling
tasks**](https://hubverse.io/en/latest/user-guide/tasks.html) that define the
expected content of a model submission against one or more modeling targets.
Each modeling task includes three properties:

 - `task_ids`: a collection of variables that can be used for modeling efforts.
   **Each unique combination of task ID values represents a single modeling task**.
 - `output_type`: the expected model output representation
 - `target_metadata`: characteristics for each target (an example of a modeling
   target is "Daily incident Flu Hospitalizations")

## Creation of a `tasks.json` file

To create the `tasks.json` file, we will start from the inside out. This means
that we will create the `target_metadata`, `task_ids`, `output_type`s first.
We will then use those to create two `model_task` objects and then bundle them
into two `round` objects, which will be inserted into a `config` object.

### Creating the `target_metadata` objects

The `target_metadata` provides both human-readable (`target_name` and
`target_units`) and machine-readable (e.g. `target_type` and `is_step_ahead`)
information about the targets.

```{r create-target-metadata}
library(hubAdmin)

target_metadata_hosp <- create_target_metadata_item(
  target = "inc hosp",
  target_name = "Weekly incident influenza hospitalizations",
  target_units = "rate per 100,000 population",
  target_keys = list(target = "inc hosp"),
  target_type = "discrete",
  is_step_ahead = TRUE,
  time_unit = "week"
)

target_metadata_death <- create_target_metadata_item(
  target = "inc death",
  target_name = "Weekly incident influenza deaths",
  target_units = "rate per 100,000 population",
  target_keys = list(target = "inc death"),
  target_type = "discrete",
  is_step_ahead = TRUE,
  time_unit = "week"
)

target_metadata <- create_target_metadata(target_metadata_hosp, target_metadata_death)
```

If you inspect the `target_metadata_hosp` object, you will see that it appears
as a class `target_metadata_item` with additional attributes about the schema
that created it:

```{r tmh}
str(target_metadata_hosp)
```


Likewise, `target_metadata` is a combination of `target_metadata_hosp`,
and `target_metadata_death`:


```{r, tme}
str(target_metadata)
```


If this process just creates lists with metadata, then why bother using `hubAdmin` functions at all to create them?
The benefit is that `hubAdmin` `create_*()` functions provide some basic validation of objects when creating them, helping you catch some potential mistakes sooner. For
example, some of the values are interdependent and if you accidentally leave one
out, the function will provide a helpful error:

```{r tmh-example}
#| error: true
create_target_metadata_item(
  target = "inc hosp",
  target_name = "Weekly incident influenza hospitalizations",
  target_units = "rate per 100,000 population",
  target_keys = list(target = "inc hosp"),
  target_type = "discrete",
  is_step_ahead = TRUE
)
```

Now that we have defined the targets in `target_metadata`, we can move on
to defining the task ID objects, which will define what the modeling parameters
will be.

### Creating the `task_id` objects

In this modeling effort, we want to **require** modelers to provide week-ahead
predictions for the "weekly incident influenza hospitalizations" target
(`inc hosp`). Model submissions can optionally include five states, provide up
to 4-week-ahead predictions, and also provide predictions for the "weekly
incident influenza deaths" target (`inc death`).

```{r set-task-ids}
origin_date <- create_task_id(
  "origin_date",
  required = NULL,
  optional = c("2023-01-02", "2023-01-09", "2023-01-16")
)

location <- create_task_id(
  "location",
  required = "US",
  optional = c("01", "02", "04", "05", "06")
)

horizon <- create_task_id("horizon", required = 1L, optional = 2:4)

target <- create_task_id("target", required = "inc hosp", optional = "inc death")

task_ids_example <- create_task_ids(origin_date, location, horizon, target)

```

Now that we've created the task IDs that specify the modeling tasks, we can define the outputs we
expect from the models. 

### Creating the `output_type` objects

For this example, we want to have three output type objects: 

1. a required `"mean"` output type  
2. a required `"quantile"` output type  
3. an optional `"median"` output type 

Additionally, our two targets will accept a different combination of output
types. Specifically, target `"inc hosp"` will only accept `"mean"` and
`"median"` output types while `"inc death"` will only accept `"mean"` and
`"quantile"` output types. 

```{r create-output-type-ids}
mean_out_type <- create_output_type_mean(
  is_required = TRUE,
  value_type = "double",
  value_minimum = 0
)

median_out_type <- create_output_type_median(
  is_required = FALSE,
  value_type = "integer",
  value_minimum = 0L

)

quantile_out_type <- create_output_type_quantile(
  required = c(0.25, 0.5, 0.75),
  is_required = TRUE,
  value_type = "double",
  value_minimum = 0
)
```

Now that we have our base `output_type_item` class objects, we can combine them to create `output_type` class objects that we want to use for each particular target. 

```{r combine-output-types}
output_type_mean_median <- create_output_type(mean_out_type, median_out_type)

output_type_mean_quantile <- create_output_type(mean_out_type, quantile_out_type)
```

### Creating the `model_task` objects

As previously discussed, we want to require a "mean" output for both tasks, but we want a "quantile"
output type for the `inc death` target and an optional "median" for the `inc
hosp` target.

```{r create-model-task}
model_task_hosp <- create_model_task(
  task_ids = task_ids_example,
  output_type = output_type_mean_median,
  target_metadata = target_metadata
)

model_task_death <- create_model_task(
  task_ids = task_ids_example,
  output_type = output_type_mean_quantile,
  target_metadata = target_metadata
)

model_task_example <- create_model_tasks(model_task_hosp, model_task_death)
```
Now that we have a set of model tasks, we can create rounds for them.

### Creating the `round` objects

While the model tasks define what a model output submission should look like,
the round additionally defines the schedule of submissions and the timeframe under which a
submission is considered valid.

The most common way to create rounds is to create a single `round` object that
and specify a date task ID whose values will act as the round IDs for each round. These round IDs are used to calculate the submission window for each modeling round. In our example, we specify `"origin_date"` as the source variable for round IDs. We also specify that the submission window for a given round is a period of one week, where the beginning of the window is
4 days before the origin date and the end of the submission window is 2 days
after the origin date. This means if you have an origin date of 
`r format(as.Date("2023-01-02"), "%A %Y-%m-%d")`, then valid submission dates
for that round are from 
`r paste(format(as.Date("2023-01-02") + c(-4, 2), "%A %Y-%m-%d"), collapse = " to ")`

```{r create-rounds}
round1 <- create_round(
  round_id_from_variable = TRUE,
  round_id = "origin_date",
  round_name = "Round 1",
  model_tasks = model_task_example,
  submissions_due = list(
    relative_to = "origin_date",
    start = -4L,
    end = 2L
  )
)

rounds <- create_rounds(round1)
```

### Creating and saving the `tasks` config file

And the penultimate step is to create a config file and write it out to your
hub.

```{r create-tasks, eval = FALSE}
config <- create_config(rounds)

write_config(config)
```
```{r really-write}
#| echo: false
tmp <- tempfile(fileext = ".json")
config <- create_config(rounds, output_type_id_datatype = "auto")
write_config(config, config_path = tmp, overwrite = TRUE)
```

<details>
<summary>contents of `hub-config/tasks.json`</summary>


````json

```{r show-config}
#| echo: false
#| results: asis
writeLines(readLines(tmp))
```

````

</details>


## Validating your config file

One of the strengths of specifying configuration in a JSON format is that it's
machine-readable, which means that no matter how complex a JSON file gets, **it
can be easily and rapidly validated**. All hubverse configuration files are
validated against a central set of [hubverse
schemas](https://github.com/hubverse-org/schemas/) that specify _how_ a
configuration file is constructed. This allows any tool to be able to read a
JSON file and validate it against the schema to ensure it's valid. 

To check that your configuration file is valid, you can use the
`validate_config()` function from the root directory of your hub:

```{r validate-config-show}
#| eval: false
validate_config(config = "tasks")
```

```{r validate_config}
#| echo: false
res <- validate_config(config_path = tmp)
attr(res, "config_path") <- "/path/to/hub/hub-config/tasks.json"
res
```

### The importance of validation

Validating your config files is important because while we can validate simple
relationships during the construction of the configuration file, it's not always
possible to validate the relationships between the elements until the final
configuration file is written and can be validated against the schema.

Take for example if we wanted to create two rounds to our current config that
has a different submission window from `"2023-01-09"` to "`2023-01-16"` (note that this
round contains an error which we will explain later):

```{r round2-fight}
round2 <- create_round(
  round_id_from_variable = TRUE,
  round_id = "origin_date",
  round_name = "Round 2",
  model_tasks = model_task_example,
  submissions_due = list(start = "2023-01-09", end = "2023-01-16")
)
two_rounds <- create_rounds(round1, round2)
new_config <- create_config(two_rounds)
```

```{r create-tasks-too, eval = FALSE}
write_config(new_config, overwrite = TRUE)
```
```{r really-write-too}
#| echo: false
tmp2 <- tempfile(fileext = ".json")
write_config(new_config, config_path = tmp2, overwrite = TRUE)
```

Now we can attempt to run a validation:

```{r validate-config-show-too}
#| eval: false
validation_result <- validate_config(config = "tasks")
validation_result
```

```{r validate_config-too}
#| echo: false
validation_result <- validate_config(config_path = tmp2)
attr(validation_result, "config_path") <- "/path/to/hub/hub-config/tasks.json"
validation_result
```


Now we are getting a notification about six errors. Let's look at the error
table: 


```{r err}
#| results: asis
view_config_val_errors(validation_result)
```

The reason why this came up is because we were defining two rounds that shared
identifiers because they both inherited identifiers from `"origin_date"`. Each
round must have a separate and unique identifier. This is because we forgot to
set `round_id_from_variable = FALSE` in our second round. If we change that,
the config becomes valid:

```{r round2-agree}
round2 <- create_round(
  round_id_from_variable = FALSE, # do not set the round ID from a task ID
  round_id = "round-2", # use the round ID of "round-2"
  round_name = "Round 2",
  model_tasks = model_task_example,
  submissions_due = list(start = "2023-01-09", end = "2023-01-16")
)
two_rounds <- create_rounds(round1, round2)
new_config <- create_config(two_rounds, output_type_id_datatype = "auto")
```

```{r create-tasks-agree, eval = FALSE}
write_config(new_config, overwrite = TRUE)
```
```{r really-write-agree}
#| echo: false
tmp2 <- tempfile(fileext = ".json")
write_config(new_config, config_path = tmp2, overwrite = TRUE)
```

Now we can attempt to run a validation:

```{r validate-config-show-agree}
#| eval: false
validation_result <- validate_config(config = "tasks")
validation_result
```

```{r validate_config-agree}
#| echo: false
validation_result <- validate_config(config_path = tmp2)
attr(validation_result, "config_path") <- "/path/to/hub/hub-config/tasks.json"
validation_result
```



